Two Site, Single Disturbance, MultiPool

### 2021-06-14 TwoSite_SingleDisturb_MultiPool.Rmd
Results: exp 5 two site plot

Na <- NaK <- abind::abind(Na.orig,
                          Na.orig[sample.int(ne,ne),],  ## resample IC for site 2
                          along = 3)
NfK = Nf = Na*NA
THETAK <- THETA
d.prior = matrix(nrow=ne,ncol=2)
muf <- matrix(NA,3*nsite,4)
pf  <- array(0,c(3*nsite,3*nsite,4)) ## 3rd variable is class
for(t in 1:10){

@@

  ## Disturbance
  for(i in 1:ne) d.prior[i,] <- rcbern(rep(p.dist,2),rho=0.5)   ## prior on disturbance status
  d.class <- d.prior[,1] + 2*d.prior[,2]
  for(d in 2:4){  ## loop over disturbance classes
    sel.dist <- which(d.class == d-1)              ## select disturbed ensemble members
    if(length(sel.dist)> (3*nsite)){
      for(j in sel.dist){                      ## run disturbance model
        switch (as.character(d),
            "2" = {Nf[j,,1] = disturbance(Nf[j,,1])
                   NfK[j,,1]= disturbance(NfK[j,,1])
                  },
            "3" = {Nf[j,,2] = disturbance(Nf[j,,2])
                   NfK[j,,2]= disturbance(NfK[j,,2])
                  },
            "4" = {Nf[j,,1] = disturbance(Nf[j,,1])
                   Nf[j,,2] = disturbance(Nf[j,,2])
                   NfK[j,,1]= disturbance(NfK[j,,1])
                   NfK[j,,2]= disturbance(NfK[j,,2])
                  }
        )
      }
      muf[,d]  = apply(Nf[sel.dist,,],3,colMeans)
      pfTMP    = try(solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2]))))
    } 
    if(sum(sel.dist) <= (3*nsite) || class(pfTMP) == "try-error") { 
      ## set to highly unlikely values that will be selected out
      muf[,d]  = rep(0,nsite*3)
      pfTMP    = solve(diag(rep(0.01,nsite*3)^2))
    }
    pf[,,d]    = pfTMP
  }
  ## set undisturbed
  sel.dist = which(d.class == 0)
  muf[,1] = apply(Nf[sel.dist,,],3,colMeans)
  pf[,,1] = solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2])))
  Nf <- tobit(Nf)

  ## EnKF forecast stats
  NfK <- tobit(NfK)
  NK = cbind(NfK[,,1],NfK[,,2])
  mufK = colMeans(NK)
  PfK  = var(NK)
  pfK  = solve(PfK)             ## Not disturbed forecast precision
  Forecast.KD[[t]] <- list(d.prior=d.prior,Nf=NfK,muf=mufK,pf=pfK)

  ## priors
  priors <- list(N = 4,  ## 2 sites x 2 disturbance classes
                 p   = table(d.class)/ne,         ## disturbance rate
                 muf = muf,    ## forecast means
                 pf  = pf      ## forecast precision
  )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## Analysis
  update = c(list(Y=No[t,],R=solve(R)),priors)
  mod <- rjags::jags.model(file=textConnection(Analysis2SMDMP),
                           data=update,
                           n.adapt=1000,n.chains=3)
  jdat <- rjags::coda.samples(mod,variable.names="X",n.iter=2000) ## burn-in 
  jdat <- rjags::coda.samples(mod,variable.names=c("X","rho","x"),n.iter=10000)

@@

  ## EnKF Analysis
  I = H <- diag(rep(1,3*nsite))
  K <- PfK %*% t(H) %*% solve(H%*%PfK%*%t(H) + R)
  muaK <- mufK + K%*%(No[t,] - mufK) 
  PaK <- (I-K %*% H)%*%PfK    
  paK = solve(PaK)
  NaK <- PEcAn.assim.sequential::adj.ens(Pf = PfK,X = NK,mu.f = mufK,mu.a = muaK,Pa = PaK) %>% as.matrix()
  Analysis.KD[[t]] <- list(mua=muaK,pa=paK,Pa=PaK,Na=NaK)
  NaK <- abind::abind(NaK[,1:3],NaK[,4:6],along = 3)

@@

>>>>

Na <- NaK <- abind::abind(Na.orig,
                          Na.orig[sample.int(ne,ne),],  ## resample IC for site 2
                          along = 3)
NfK = Nf = Na*NA
THETAK <- THETA
d.prior = matrix(nrow=ne,ncol=2)
muf <- matrix(NA,3*nsite,4)
pf  <- array(0,c(3*nsite,3*nsite,4)) ## 3rd variable is class
for(t in 1:NT){

@@

  ## Disturbance
  for(i in 1:ne) d.prior[i,] <- rcbern(rep(p.dist,2),rho=0.5,mu=1:2)   ## prior on disturbance status
  d.class <- d.prior[,1] + 2*d.prior[,2]
  for(d in 2:4){  ## loop over disturbance classes
    sel.dist <- which(d.class == d-1)              ## select disturbed ensemble members
    if(length(sel.dist)> (3*nsite)){
      for(j in sel.dist){                      ## run disturbance model
        switch (as.character(d),
                "2" = {Nf[j,,1] = disturbance(Nf[j,,1])
                NfK[j,,1]= disturbance(NfK[j,,1])
                },
                "3" = {Nf[j,,2] = disturbance(Nf[j,,2])
                NfK[j,,2]= disturbance(NfK[j,,2])
                },
                "4" = {Nf[j,,1] = disturbance(Nf[j,,1])
                Nf[j,,2] = disturbance(Nf[j,,2])
                NfK[j,,1]= disturbance(NfK[j,,1])
                NfK[j,,2]= disturbance(NfK[j,,2])
                }
        )
      }
      muf[,d]  = apply(Nf[sel.dist,,],3,colMeans)
      pfTMP    = try(solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2]))))
    } 
    if(sum(sel.dist) <= (3*nsite) || class(pfTMP) == "try-error") { 
      ## set to highly unlikely values that will be selected out
      muf[,d]  = rep(0,nsite*3)
      pfTMP    = solve(diag(rep(0.01,nsite*3)^2))
    }
    pf[,,d]    = pfTMP
  }
  ## set undisturbed
  sel.dist = which(d.class == 0)
  muf[,1] = apply(Nf[sel.dist,,],3,colMeans)
  pf[,,1] = solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2])))
  Nf <- tobit(Nf)

  ## EnKF forecast stats
  if(FALSE){
    NfK <- tobit(NfK)
    NK = cbind(NfK[,,1],NfK[,,2])
    mufK = colMeans(NK)
    PfK  = var(NK)
    pfK  = solve(PfK)             ## Not disturbed forecast precision
    Forecast.KD[[t]] <- list(d.prior=d.prior,Nf=NfK,muf=mufK,pf=pfK)
  }
  ## priors
  priors <- list(N = 4,  ## 2 sites x 2 disturbance classes
                 p   = table(d.class)/ne,         ## disturbance rate
                 muf = muf,    ## forecast means
                 pf  = pf      ## forecast precision
  )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## initial conditions: disturbance likelihood
  Lp = NA
  for(i in 1:4){ Lp[i] = mvtnorm::dmvnorm(No[t,],muf[,i],pf[,,i])}
  rho.IC = rep(0,4); rho.IC[which.max(Lp)] = 1; rho.IC[1] = NA

  ## Analysis
  update = c(list(Y=No[t,],R=solve(R)),priors)
  mod <- rjags::jags.model(file=textConnection(Analysis2SMDMP),
                           data=update,
                           inits = list(rho=rho.IC),
                           n.adapt=1000,n.chains=3)
  jdat <- rjags::coda.samples(mod,variable.names="rho",n.iter=2000) ## burn-in 
  jdat <- rjags::coda.samples(mod,variable.names=c("X","rho","x"),n.iter=10000)

@@

  ## EnKF Analysis
  if(FALSE){
    I = H <- diag(rep(1,3*nsite))
    K <- PfK %*% t(H) %*% solve(H%*%PfK%*%t(H) + R)
    muaK <- mufK + K%*%(No[t,] - mufK) 
    PaK <- (I-K %*% H)%*%PfK    
    paK = solve(PaK)
    NaK <- PEcAn.assim.sequential::adj.ens(Pf = PfK,X = NK,mu.f = mufK,mu.a = muaK,Pa = PaK) %>% as.matrix()
    Analysis.KD[[t]] <- list(mua=muaK,pa=paK,Pa=PaK,Na=NaK)
    NaK <- abind::abind(NaK[,1:3],NaK[,4:6],along = 3)
  }

@@

```{r}
save(Forecast,Analysis,No,Ntrue,R,file="exp5.RData")
Forecast10 = Forecast[[10]]
Analysis10 = Analysis[[10]]
CI.Na = list()
for(i in c(3,6)){
  CI.Na[[i]] = sapply(X = Analysis,function(y){
    x = y$MCMC
    sel.n = which(colnames(x) == paste0("X[",i,"]"))
    quantile(x[,sel.n],c(0.025,0.5,0.975))
  })
}
save(Forecast10,Analysis10,CI.Na,No,Ntrue,R,file="exp5trim.RData")

### 2021-06-14 results.Rmd
Results: exp 5 two site plot

## ensemble adjustment
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)
```

@@

```{r, echo=TRUE,fig.cap="Simulated experiments 3-5: Left hand column shows timeseries for the wood pool for all three experiments; Right hand shows bivariate ensemble adjustment at time of disturbance"}
## figure with 3 rows, 2 col showing timeseries in rt and bivariate in left

@@

## ensemble adjustment
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)

@@

## ensemble adjustment
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)

#experiment 5 - one disturb, 2 sites
## Generated by

```

>>>>

## ensemble adjustment
ne = nrow(Forecast[[10]]$Nf)
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)
```

@@

```{r, echo=FALSE,fig.cap="Simulated experiments 3-5: Left hand column shows timeseries for the wood pool for all three experiments; Right hand shows bivariate ensemble adjustment at time of disturbance"}
## figure with 3 rows, 2 col showing timeseries in rt and bivariate in left

@@

## ensemble adjustment
ne = nrow(Forecast[[10]]$Nf)
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)

@@

## ensemble adjustment
ne = nrow(Forecast[[10]]$Nf)
sel.a= seq(1,ne,length=10)#sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1.5)

######### experiment 5 - one disturb, 2 sites
load("exp5trim.Rdata") ##Generated by TwoSite_SingleDisturb_MultiPool.Rmd
pool.names=c("Cleaf","Csoil","Cwood")
rng = range(CI.Na)
if(rng[1]<1) rng[1] = 0
plot(No[,i],
     ylim = rng,type='n',
     ylab=pool.names[i],xlab="time",
     main="Multi-Site, Single-Disturbance")
  ecoforecastR::ciEnvelope(1:ncol(CI.Na[[3]]),CI.Na[[3]][1,],CI.Na[[3]][3,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  ecoforecastR::ciEnvelope(1:ncol(CI.Na[[6]]),CI.Na[[6]][1,],CI.Na[[6]][3,],col=ecoforecastR::col.alpha("green",0.7))
lines(CI.Na[[3]][2,],col="#b2df8a",lwd=2)
lines(CI.Na[[6]][2,],col="green",lwd=2)
points(No[,3])
points(No[,6],pch=2)
legend("bottomleft",legend=c("Site 1","Site 2","pseudodata"),
       lwd=5,col=c("#b2df8a","green","black"))

### cov plot
ss = floor(seq(1,nrow(Analysis10$Na),length=200)) ## subsample
plot(Analysis10$Na[,1],Analysis10$Na[,3],pch=Analysis10$Ca,col=ecoforecastR::col.alpha("#b2df8a",alpha=1),
     ylim=range(Forecast10$Nf[,3,1]),
     xlim=range(Forecast10$Nf[,1,1]),
     xlab=pool.names[1],ylab=pool.names[3],
     cex.lab=1.25,asp = 1,type='n')
points(Forecast10$Nf[ss,1,1],Forecast10$Nf[ss,3,1],cex=0.5,col="purple",pch=18)
points(Forecast10$Nf[ss,1,2],Forecast10$Nf[ss,3,2],cex=0.5,col="violet",pch=18)
points(Analysis10$Na[ss,1],Analysis10$Na[ss,3],pch=Analysis10$Ca[ss],col=ecoforecastR::col.alpha("#b2df8a",alpha=0.8),cex=0.5)
points(Analysis10$Na[ss,4],Analysis10$Na[ss,6],pch=Analysis10$Ca[ss],col=ecoforecastR::col.alpha("green",alpha=0.8),cex=0.5)
points(No[10,1],No[10,3],cex=3,col=1)
lines(No[10,1]+c(-1,1)*sqrt(R[1,1]),rep(No[10,3],2))
lines(rep(No[10,1],2),No[10,3]+c(-1,1)*sqrt(R[3,3]))
points(No[10,4],No[10,6],cex=3,col=1)
lines(No[10,4]+c(-1,1)*sqrt(R[1,1]),rep(No[10,6],2))
lines(rep(No[10,4],2),No[10,6]+c(-1,1)*sqrt(R[3,3]))

## ensemble adjustment
sel.a= ss[seq(1,length(ss),length=10)]#sample.int(ne,10)
arrows(Forecast10$Nf[sel.a,1,1],Forecast10$Nf[sel.a,3,1],
       Analysis10$Na[sel.a,1],Analysis10$Na[sel.a,3],length = 0.1,lwd = 0.5)
arrows(Forecast10$Nf[sel.a,1,2],Forecast10$Nf[sel.a,3,2],
       Analysis10$Na[sel.a,4],Analysis10$Na[sel.a,6],length = 0.1,lwd = 0.5)
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c("purple","#b2df8a",1),
       cex = 1)
```

## Remotely-sensed disturbance in Oregon
